---
title: "eugenie_derrien"
author: "Eugénie Derrien"
format: html
---

## Types and classes 

### Types 
```{r}
typeof(1.5)
typeof("toto")
typeof('also between single quotes')
typeof(TRUE)
typeof(1L)
typeof(1)
typeof(typeof)
typeof(list())
```
What is a type ? In computer science, the way you can manipulate objects depends on objects type. If I manipulate a double (real numbers), I excepts the computer to provide me with usual operations on real numbers. 

Different types : 

-double -> real numbers
-integers -> integers when specified (if not specified it will be a double)
-character -> need to use libraries in R to do interesting with it
-logical -> TRUE or FALSE
-closure -> functions 
-list 


### Classes

Classes are your own types that you created on your computer, Types are your environment on your computer. 

```{r}
data(iris)
typeof(iris)
class(iris)
```
You can create a class from a list or a vector. 
What we have to understand is that we have operations that are available for types and that can be applied to classes. 

