---
title: "robertocanu"
author: 'Roberto Canu'
format: html
editor: visual
---

Notes for Data Management.

## INDEXING

```{r}
x <- c(6,4,7,12,21,0)
x[3]
```

```{r}
x[5] <- 89
```

The following sets y to be not the same thing but basically a copy of x:
```{r}
y <- x
```

```{r}
x[2] <- 33
```
x
y

Modifying x doesn't impact y (unlike in python and other programming languages). 

It is called a lazy copy: saying x and y are the same (y <- x) they actually are the same. But when I call the modification of one of the elements of either of them, a copy is created for the purpose. 

```{r}
x
x[c(2,2,1,4,5)]
```
This puts the number associated with each of the position written in c from x itself. 

```{r}
x[0]
```
Returns an empty numeric vector because index 0 has no meaning in R. 

```{r}
typeof(NA)
length(NA)
```

```{r}
typeof(NA_integer_)
```
Now the type is integer.

```{r}
x
x[c(4,1,0,15,6)]
```
`0` gets disregarded entirely, so no respective value is found in the new vector; `15` is not a position in x (that has 6 elements) and thus `NA` is printed instead. 

```{r}
y <- c('A', 'B', 'C')
y[c(TRUE, FALSE, FALSE)]
```

```{r}
x
x[c(-1,-4,-5)]
x[c(-2,-3,-6)]
```
We must use either positive indexes or negative ones. The negative indexes remove the value at that position.


```{r}
x
x[c(2,3,10)] <- c(7,88,190)
x
```
The vector gets enlarged so that he has the position needed (the biggest one required, in this case 10), and filled with `NAs` in the other missing positions. 

